---
output:
  pdf_document: default
  html_document: default
---

CONCRETE COMPRESSIVE STRENGTH


Load the required Packages
```{r}
library(tidyverse)
library(corrplot)
library(readxl)
library(datarium)
library(qqplotr)
library(rcompanion)
library(lattice)
library(car)
library(FSA)
library(multcomp)
library(RVAideMemoire)
```

Load the data
```{r}
df <- read_excel('concrete compressive strength.xlsx')
df <- as.data.frame(df)
# check the head of the dataset
head(df)
```
TASK 1 - EDA AND SUMMARY STATISTICS

```{r}
# check for missing values
any(is.na(df))
# check for duplicates
sum(duplicated(df))
```
No missing values, but there are 25 duplicate rows, these duplicates would be removed

```{r}
# remove duplicate rows
df <- df %>% distinct()
```


```{r}
str(df)
```


To make things easier, we are going to rename to columns


```{r}
names(df)
```


```{r}
# Rename the columns to easier names
df <- df %>% 
  rename(
    cement = `Cement (component 1)(kg in a m^3 mixture)`,
    blast_furnace_slag = `Blast Furnace Slag (component 2)(kg in a m^3 mixture)`,
    fly_ash = `Fly Ash (component 3)(kg in a m^3 mixture)`,
    water = `Water  (component 4)(kg in a m^3 mixture)`,
    superplasticizer = `Superplasticizer (component 5)(kg in a m^3 mixture)`,
    coarse_aggregate = `Coarse Aggregate  (component 6)(kg in a m^3 mixture)`,
    fine_aggregate = `Fine Aggregate (component 7)(kg in a m^3 mixture)`,
    age = `Age (day)`,
    concrete_category = `Concrete Category`,
    contains_fly_ash = `Contains Fly Ash`,
    strength = `Concrete compressive strength(MPa, megapascals)`
  )

names(df)
```
```{r}
#check the structure of the data
str(df)
```

We can see that we have 11 columns, 9 are numerical, 2 categorical 


Summary Statistics

```{r}
#check summary of the data to get a quick view of the means, medians, ranges, etc.
summary(df)
```
With just quick glance, we can see that some of the columns have a widely different means and medians, 
e.g Blast Furnace Slag, Fly ash, etc. This means that these columns are skewed! more EDA to verify.



Distribution of the columns and Tests for Normality

We set alpha = 0.05


A. Histogram & Density plot for Numerical columns, and Tests for Normality
```{r}
# Plot a histogram, and add a density plot to show the distribution
df %>% 
  ggplot(aes(x=cement))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Cement", x = "Cement (kg/m3)")+
  theme_bw()
```
Cement: Slightly right-skewed distribution

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=cement))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Cement", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Cement: Deviations from the line suggest non-normality, particularly in the tails.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$cement)
```
p-value < 0.05, Not normal


```{r}
df %>% 
  ggplot(aes(x=blast_furnace_slag))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Blast Furnace Slag", x = 'Blast Furnace Slag (kg/m3)')+
  theme_bw()
```
Blast Furnace Slag: Strong right skew, with many zero values.


```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=blast_furnace_slag))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Blast Furnace Slag", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Blast Furnace Slag: Significant deviations from normality, confirming the right-skewed nature.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$blast_furnace_slag)
```
p-value < 0.05, Not normal

```{r}
df %>% 
  ggplot(aes(x=fly_ash))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Fly Ash", x = "Fly Ash (kg/m3)")+
  theme_bw()
```
Fly Ash: Similar to blast furnace slag, right-skewed with many zero values.

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=fly_ash))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Fly Ash", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Fly Ash: Significant deviations from normality, confirming the right-skewed nature.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$fly_ash)

```
p-value < 0.05, Not normal

```{r}
df %>% 
  ggplot(aes(x=water))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Water", x = 'Water (kg/m3)')+
  theme_bw()
```
Water: More symmetrical distribution, with values mostly centered around the mean, suggesting less skew.

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=water))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Water", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Water: Close to normal, but slight deviations in the tails.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$water)

```

```{r}
df %>% 
  ggplot(aes(x=superplasticizer))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Superplasticizer", x = 'Superplasticizer (kg/m3)')+
  theme_bw()
```
Superplasticizer: Highly skewed, with many zero values.

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=superplasticizer))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Superplasticizer", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Superplasticizer: Deviates from the line, confirming non-normality.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$superplasticizer)

```


```{r}
df %>% 
  ggplot(aes(x=coarse_aggregate))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Coarse Aggregate", x = 'Coarse Aggregate (kg/m3)')+
  theme_bw()
```
Coarse Aggregate: Roughly symmetric

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=coarse_aggregate))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Coarse Aggregate", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Coarse Aggregate: slight deviations, probably none normal.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$coarse_aggregate)

```


```{r}
df %>% 
  ggplot(aes(x=fine_aggregate))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Fine Aggregate", x = 'Fine Aggregate (kg/m3)')+
  theme_bw()
```
Fine Aggregate: Nearly normal, although slightly skewed to the left.

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=fine_aggregate))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Fine Aggregate", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Fine Aggregate:slight deviations, probably none normal.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$fine_aggregate)

```



```{r}
df %>% 
  ggplot(aes(x=age))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Age", x = 'Age (Days)')+
  theme_bw()
```
Age: Right-skewed, many samples at lower ages.

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=age))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Age", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Age: Major deviations from the line, confirming non-normality.

```{r}
# Shapiro-Wilk Test
shapiro.test(df$age)

```

```{r}
df %>% 
  ggplot(aes(x=strength))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, fill = 'steelblue', alpha = 0.5, color = 'black')+
  geom_density(color = 'darkred', linewidth=0.75)+
  labs(title = "Distribution of Concrete Compressive Strength", x = 'Concrete Compressive Strength (Mpa)')+
  theme_bw()
```
Compressive Strength: Somewhat symmetric but with slight right skew.

```{r}
# QQ plot
df %>% 
  ggplot(aes(sample=strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Concrete Compressive Strength", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Compressive Strength: Minor deviations, could be normal?

```{r}
# Shapiro-Wilk Test
shapiro.test(df$strength)

```

All p-values are less than 0.05, so we reject the null hypothesis of normality, confirming that all the continuous columns are non-normal.


PLOT SCATTER PLOTS VS COMPRESSIVE STRENGTHS FOR NUMERIC COLUMNS
```{r}
df %>% 
  ggplot(aes(x=cement, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "Cement vs Compressive Strength Grouped by Category \n& Presense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```
Cement vs. Compressive Strength: Positive trend across categories, indicating that higher cement content generally increases compressive strength

```{r}
df %>% 
  ggplot(aes(x=blast_furnace_slag, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "Blast Furnace Slag vs Compressive Strength Grouped by Category & \nPresense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```
Blast Furnace Slag: somewhat positive relationship with compressive strength.

```{r}
df %>% 
  ggplot(aes(x=fly_ash, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "Fly Ash vs Compressive Strength Grouped by Category \n& Presense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```

Fly Ash: Multiple observations without fly ash, inverse relationship with concrete strength.

```{r}
df %>% 
  ggplot(aes(x=water, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "Water vs Compressive Strength Grouped by Category \n& Presense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```
Water: Inverse relationship, suggesting higher water content may reduce compressive strength.

```{r}
df %>% 
  ggplot(aes(x=superplasticizer, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "Superplasticizer vs Compressive Strength Grouped by Category \n& Presense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```
Superplasticizer: Positive association with compressive strength in most cases.

```{r}
df %>% 
  ggplot(aes(x=coarse_aggregate, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "Coarse Aggregate vs Compressive Strength Grouped by Category \n& Presense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```
Coarse Aggregate: Coarse aggregate shows slightly negative association

```{r}
df %>% 
  ggplot(aes(x=fine_aggregate, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "FIne Aggregate vs Compressive Strength Grouped by Category & \nPresense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```
Fine Aggregate: Coarse aggregate shows generally negative association


```{r}
df %>% 
  ggplot(aes(x=age, y=strength, color=concrete_category))+
  geom_point(alpha=0.7)+
  geom_smooth(color='black', method = "lm")+
  labs(title = "Age vs Compressive Strength Grouped by Category \n& Presense of Fly Ash",
       y="Compressive Strength (MPa)")+
  facet_grid(concrete_category ~ contains_fly_ash)+
  theme_bw()
```
Age: Positive relationship with compressive strength, indicating concrete strengthens over time.


BOX PLOTS TO CHECK FOR OUTLIERS
```{r}
df %>%
  ggplot(aes(y=cement))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of cement', y = "Cement (kg/m3)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=blast_furnace_slag))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of Blast Furnace Slag', y = "Blast Furnace Slag (kg/m3)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=fly_ash))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of Fly ash', y = "Fly Ash (kg/m3)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=water))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of Water', y = "Water (kg/m3)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=superplasticizer))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of Superplasticizer', y = "Superplasticizer (kg/m3)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=coarse_aggregate))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of Coarse aggregate', y = "Coarse Aggregate (kg/m3)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=fine_aggregate))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of Fine aggregate', y = "Fine Aggregate (kg/m3)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=age))+
  geom_boxplot(fill='steelblue', color='black')+
  labs(title = 'Box plot of Age', y = "Age (days)")+
  theme_bw()
```


```{r}
df %>%
  ggplot(aes(y=strength, x=concrete_category, fill=concrete_category))+
  geom_boxplot(color='black')+
  labs(title = 'Box plot of Concrete Strength by Concrete Category \nand Fly Ash Presence', 
       y='Compressive Strength (MPa)')+
  facet_grid(~contains_fly_ash)+
  theme_bw()
```



B - Categorical Variables
```{r}
df %>% 
  ggplot(aes(x=concrete_category))+
  geom_bar(fill = 'steelblue', alpha = 0.5, color = 'black')+
  labs(title = "Count of Concrete Category", x = 'Concrete Category')+
  theme_bw()
```


```{r}
df %>% 
  ggplot(aes(x=contains_fly_ash))+
  geom_bar(fill = 'steelblue', alpha = 0.5, color = 'black')+
  labs(title = "Count of Contains Fly Ash", x = 'Contains Fly Ash')+
  theme_bw()
```
BOX PLOT
```{r}
df %>% 
  ggplot(aes(x=concrete_category, y=strength))+
  geom_boxplot(color='black', fill='steelblue')+
  labs(title='Compressive Strength by Concrete Category', y='Compressive Strength (MPa)')+
  theme_bw()
```

```{r}
df %>% 
  ggplot(aes(x=contains_fly_ash, y=strength))+
  geom_boxplot(color='black', fill='steelblue')+
  labs(title='Compressive Strength by Contains Fly Ash', y='Compressive Strength (MPa)')+
  theme_bw()
```


CORRELATION ANALYSIS

A. Correlation Matrix for Continuous Variables
```{r}
# Select all the continuous variables
continous_vars <- df %>% 
  dplyr::select(cement, blast_furnace_slag, fly_ash, water, superplasticizer, 
         coarse_aggregate, fine_aggregate, age, strength)

head(continous_vars)
```


```{r}
# get the correlation matrix using spearman, and round to 3 decimal places
corr_matrix <- round(cor(continous_vars, method = 'spearman'), digits = 2)

corr_matrix
```


```{r}
# Visualize it using heatmap
corrplot(corr_matrix, method = "number", type = "upper", tl.col = "black",
         title = ("Spearman Correlation for Continuos Variables"),
         # mar to adjust the margins, cex to reduce the text size
         mar = c(1, 0, 3, 0), number.cex = 0.6)
```
Age and Strength (0.61): This moderate positive correlation suggests that as the concrete ages, its compressive strength generally increases. 

Cement and Strength (0.46): This moderate positive correlation indicates that higher cement content is associated with greater compressive strength.

Water and Superplasticizer (-0.67): This strong negative correlation suggests that as water content increases, superplasticizer content decreases, and vice versa.

Fly Ash and Cement (-0.41): This moderate negative correlation indicates that mixes with higher fly ash content tend to use less cement. 


B - Correlation between the 2 Categorical Variables (Contains Fly Ash & Concrete Category)
    Both are Nominal, so we use Cramer V
```{r}
cramerV(df$concrete_category, df$contains_fly_ash)
```
Concrete Category and Contains Fly Ash (Cramer’s V = 0.02157):
This very low Cramer’s V value implies virtually no association between concrete category (Fine vs. Coarse) and the presence of fly ash. This suggests that fly ash usage does not vary between Fine and Coarse concrete mixes.




C - Correlation Between Continuous Variables and Binary Categorical Variable
    Use Point Biseral correlation
    
```{r}
# create new columns to change the datatypes of the categorical variables to numeric (0 and 1)
df_copy <- df %>% 
  mutate(contains_fly_ash_numeric = as.numeric(contains_fly_ash),
         concrete_category = as.factor(concrete_category),  # convert it to factor first
         concrete_category_numeric = as.numeric(concrete_category) - 1) # minus 1 to make it 0 and 1

head(df_copy)
```


```{r}
# Perform point biserial correlation on cement and concrete_category
cor.test(df$cement, df_copy$concrete_category_numeric)
```
Cement and Concrete Category (-0.001889355): This near-zero correlation indicates that cement content does not differ meaningfully between Fine and Coarse concrete categories. Both categories likely use a similar range of cement proportions.


```{r}
# Perform point biserial correlation on cement and contains_fly_ash
cor.test(df$cement, df_copy$contains_fly_ash_numeric)
```
Cement and Contains Fly Ash (-0.3335427): This moderate negative correlation suggests that mixes containing fly ash tend to have lower cement content.


```{r}
# Perform point biserial correlation on strength and contains_fly_ash
cor.test(df$strength, df_copy$contains_fly_ash_numeric)
```
Strength and Contains Fly Ash (-0.03390848): This near-zero negative correlation suggests a minimal relationship between fly ash presence and compressive strength.

```{r}
# Perform point biseral correlation on strength and concrete_category
cor.test(df$strength, df_copy$concrete_category_numeric)
```
Strength and Concrete Category (0.01297848): This very low positive correlation indicates that compressive strength does not really vary by concrete category (Fine vs. Coarse).






REGRESSION

MULTIPLE LINEAR REGRESSION

Objective of the regression analysis:
We want to examine the possible linear relation between compressive strength and more than one Independent Variable

MODEL 1
Lets start with cement, blast_furnace_slag, water, and superplasticizer

Y = strength
X1 = cement
X2 = blast_furnace_slag
X3 = water
X4 = superplasticizer
```{r}
# MLR model with cement, blast_furnace_slag, water, and superplasticizer
model_1 <- lm(strength ~ cement + blast_furnace_slag + water + superplasticizer, continous_vars)

summary.lm(model_1)
```
All the coefficients and intercept are significant at 0.05 level

strength = 22.397972 + 0.086402 * cement + 0.053765 * blast_furnace_slag - 0.102732 * water + 0.598461 * superplasticizer

Adjusted R-squared = 0.4086, means that they can predict 40.9% of the total variability of compressive strength.


Check Assumptions

Assumption 1. Linearity between Strength and each variable
```{r}
pairs(continous_vars[, c(9, 1, 2, 4, 5)], lower.panel = NULL, pch = 19, cex=0.2)
```
All the variables appear to have an approximately linear relationship with strength

Assumption 2: Residuals' Independence
```{r}
# plot residuals vs fitted values
plot(model_1, 1)
```


Assumption 3: Normality of Residuals
```{r}
# plot qq plot of residuals
plot(model_1, 2)
```
The residuals are approximately normally distributed.

Assumption 4: Homescedacity (Equal Variance)
```{r}
plot(model_1, 3)
```


Assumption 5: No Multicolinearity
```{r}
vif(model_1)
```
All VIFs less than 5: No multicolinearity.


All 5 assumptions were approved, so we can confirm the equation

strength = 22.397972 + 0.086402 * cement + 0.053765 * blast_furnace_slag - 0.102732 * water + 0.598461 * superplasticizer 




MODEL 2: Cement, blast furnace slag, water, fly_ash

Y = strength
X1 = cement
X2 = blast_furnace_slag
X3 = water
X4 = fly_ash
```{r}
# MLR model with cement, blast_furnace_slag, water, and fly ash
model_2 <- lm(strength ~ cement + blast_furnace_slag + water + fly_ash, continous_vars)

summary.lm(model_2)
```
All the coefficients and intercept are significant at 0.05 level

strength = 23.841431 + 0.110835 * cement + 0.081224 * blast_furnace_slag - 0.159829 * water + 0.067997 * fly_ash

Adjusted R-squared = 0.42, means that they can predict 42% of the total variability of compressive strength. -> Best model


Check Assumptions

Assumption 1. Linearity between Strength and each variable
```{r}
pairs(continous_vars[, c(9, 1, 2, 4, 3)], lower.panel = NULL, pch = 19, cex=0.2)
```
All the variables appear to have an approximately linear relationship with strength

Assumption 2: Residuals' Independence
```{r}
# plot residuals vs fitted
plot(model_2, 1)
```
The correlation is approximately 0

Assumption 3: Normality of Residuals
```{r}
#plot qq plot of residuals
plot(model_2, 2)
```
The residuals are approximately normally distributed.

Assumption 4: Homescedacity (Equal Variance)
```{r}
plot(model_2, 3)
```
The variance is constant.

Assumption 5: No Multicolinearity
```{r}
vif(model_2)
```
All VIFs less than 5, so no multicolinearity.


ALL 5 ASSUMPTIONS WERE APPROVED, THEEFORE:

strength = 24.369098 + 0.112530 * cement + 0.084340 * blast_furnace_slag - 0.165824 * water + 0.068003 * fly_ash


MODEL 3

Y = strength
X1 = cement
X2 = water
```{r}
# MLR model with cement, blast_furnace_slag, water, and fly ash
model_3 <- lm(strength ~ cement + water, continous_vars)

summary.lm(model_3)
```
All the coefficients and intercept are significant at 0.05 level

strength = 48.34263 + 0.07406 * cement - 0.18524 * water

Adjusted R-squared = 0.31, means that they can predict 31% of the total variability of compressive strength. 


Check Assumptions

Assumption 1. Linearity between Strength and each variable
```{r}
pairs(continous_vars[, c(9, 1, 4)], lower.panel = NULL, pch = 19, cex=0.2)
```
All the variables appear to have an approximately linear relationship with strength

Assumption 2: Residuals' Independence
```{r}
# plot residuals vs fitted
plot(model_3, 1)
```
The residuals are independent

Assumption 3: Normality of Residuals
```{r}
#plot qq plot of residuals
plot(model_3, 2)
```
The residuals are approximately normally distributed.

Assumption 4: Homescedacity (Equal Variance)
```{r}
plot(model_3, 3)
```
The variance is constant

Assumption 5: No Multicolinearity
```{r}
vif(model_3)
```
VIF less than 5, so no multicolinearity

All 5 Assumptions were met, THEEFORE:

strength = 48.34263 + 0.07406 * cement - 0.18524 * water





HYPOTHESIS TESTING

1. Hypothesis Test 1: 

alpha = 0.05

Null Hypothesis (H0): The mean compressive strength of concrete with fly ash is the same as the mean compressive strength of concrete without fly ash.

Alternative Hypothesis (H1): There is a significant difference in the mean compressive strength of concrete with and without fly ash. 

This is a two tailed test.

Check assumptions:

Assumption 1. check for normality

```{r}
# Filter the samples that contain fly ash
has_fly_ash <- df %>% 
  filter(contains_fly_ash == TRUE)

has_fly_ash %>% 
  ggplot(aes(sample=strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for strength with Fly Ash", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
QQ plot seems like it is approximately normally distributed

```{r}
# Filter the samples that don't contain fly ash
no_fly_ash <- df %>% 
  filter(contains_fly_ash == FALSE)

no_fly_ash %>% 
  ggplot(aes(sample=strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for strength without Fly Ash", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
QQ plot seems like there are some deviations from normality


Since at least one of the distributions is not normal, we can transform them to get an approximately normal distribution

```{r}
# Transform the distributions using sqrt

df <- df %>% 
  mutate(sqrt_strength = sqrt(strength))

# Filter the samples that contain fly ash
has_fly_ash <- df %>% 
  filter(contains_fly_ash == TRUE)

# Filter the samples that don't contain fly ash
no_fly_ash <- df %>% 
  filter(contains_fly_ash == FALSE)
```


```{r}
has_fly_ash %>% 
  ggplot(aes(sample=sqrt_strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for sqrt(strength) with Fly Ash", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
The QQ plot seems approximately normally distributed

```{r}
no_fly_ash %>% 
  ggplot(aes(sample=sqrt_strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for sqrt(strength) without Fly Ash", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
The QQ plot seems approximately normally distributed


Check their distributions
```{r}
has_fly_ash %>% 
  ggplot(aes(x=sqrt_strength))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, alpha = 0.5, color='black', fill='steelblue')+
  geom_density(color='darkred')+
  labs(title = "Distribution of SQRT Concrete Compressive Strength (With Fly Ash)")+
  theme_bw()

no_fly_ash %>% 
  ggplot(aes(x=sqrt_strength))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, alpha = 0.5, color='black', fill='steelblue')+
  geom_density(color='darkred')+
  labs(title = "Distribution of SQRT Concrete Compressive Strength (Without Fly Ash)")+
  theme_bw()
```

Both Distributions appear approximately normal

Assumption 3: No significant outliers in each group
```{r}
# Check for outliers with box plot
df %>% 
  ggplot(aes(x=contains_fly_ash, y=sqrt_strength, fill=contains_fly_ash)) +
  geom_boxplot()+
  labs(title = "SQRT Compressive Strength by Presence of Fly Ash",
       x = "Contains Fly Ash",
       y = "SQRT Compressive Strength (MPa)") +
  theme_bw()
```
There are no outliers present

Assumption 3: Independence of observations  -- They are independent.

Assumption 4: Homogeneity of variances
```{r}
bartlett.test(sqrt_strength ~ contains_fly_ash, data=df)
```
p < 0.05, therefore the variances are not equal


All assumptions met for the Independent Two-sample T test except for the equal variances.
So we run WELCH T-Test instead

Run Welch T Test
```{r}
t.test(sqrt_strength ~ contains_fly_ash, data=df, var.equal=FALSE)
```
The p-value > 0.05, we fail to reject the null hypothesis

This indicates that the mean compressive strength of concrete with fly ash is the same as the mean compressive strength of concrete without fly ash.

The Welch Independent two sample T test results show no statistically significant difference in compressive strength between concrete mixes with and without fly ash (p = 0.8007).



2. Hypothesis Test 2:

alpha = 0.05

Null Hypothesis (H0): There is no difference in compressive strength between Fine and Coarse concrete categories.

Alternative Hypothesis (H1): There is a significant difference in compressive strength between Fine and Coarse concrete categories.

This is a two tailed test.

Check assumptions:

1. check for normality

```{r}
# Filter the samples that contain fly ash
fine <- df %>% 
  filter(concrete_category == "Fine")

fine %>% 
  ggplot(aes(sample=strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for strength with Fine concrete", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
QQ plot seems like it is not normal


```{r}
# Filter the samples that dont contain fly ash
coarse <- df %>% 
  filter(concrete_category == "Coarse")

coarse %>% 
  ggplot(aes(sample=strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for strength with Coarse Concrete", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
QQ plot seems not normal

Check for normality with transformed strength
```{r}
# check for normality in the transformed variable
fine %>% 
  ggplot(aes(sample=sqrt_strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for SQRT(strength) with Fine concrete", x = 'Theoretical', y = 'Sample')+
  theme_bw()

coarse %>% 
  ggplot(aes(sample=sqrt_strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for SQRT(strength) with Coarse Concrete", x = 'Theoretical', y = 'Sample')+
  theme_bw()
```
Both distributions are still NOT normally distributed


plot the distributions
```{r}
fine %>% 
  ggplot(aes(x=sqrt_strength))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, alpha = 0.5, color='black', fill='steelblue')+
  geom_density(color='darkred')+
  labs(title = "Distribution of SQRT Concrete Compressive Strength (With Fine Concrete)")+
  theme_bw()
```

```{r}
coarse %>% 
  ggplot(aes(x=sqrt_strength))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, alpha = 0.5, color='black', fill='steelblue')+
  geom_density(color='darkred')+
  labs(title = "Distribution of SQRT Concrete Compressive Strength (With Coarse Concrete)")+
  theme_bw()
```
Both Distributions appear approximately normally distributed

Assumption 3: No significant outliers
```{r}
# Check for outliers with box plot
df %>% 
  ggplot(aes(x=concrete_category, y=sqrt_strength, fill=concrete_category)) +
  geom_boxplot()+
  labs(title = "SQRT Compressive Strength by Concrete Category",
       x = "Concrete Category",
       y = "SQRT Compressive Strength (MPa)") +
  theme_bw()
```
outliers ??????

Assumption 3: Independence of observations  -- They are independent.

Assumption 4: Homogeneity of variances

```{r}
bartlett.test(sqrt_strength ~ concrete_category, data=df)
```
p < 0.05, therefore the variances are not equal


All assumptions met for the Independent Two-sample T test except for the equal variances.
So we run WELCH T-Test instead

Run Welch T Test
All assumptions met for Independent two sample T test

Run Welch T-Test
```{r}
t.test(sqrt_strength ~ concrete_category, data=df, var.equal=FALSE)
```
The p-value > 0.05, we fail to reject the null hypothesis

The Welch Independent two sample T test shows no statistically significant difference in compressive strength between Fine and Coarse concrete categories (p = 0.8844). 





3. Hypothesis Test 3: Difference in compressive strength across age groups

Null Hypothesis (H0): The mean compressive strength is the same across different age groups (Early Age, Standard Strength, Mature Age).

Alternative Hypothesis (H1): The mean compressive strength differs across these age groups.
```{r}
# Create the age groups column
df <- df %>% 
  mutate(age_group = case_when(
    age <= 7 ~ "Early",
    age >= 8 & age <= 28 ~ "Middle",
    age >= 29 ~ "Mature"
  ),
  age_group = factor(age_group, levels = c("Early", "Middle", "Mature"), ordered = TRUE))

head(df)
```
One way ANOVA ASSUMPTIONS
1. Dependent variable should be continuous: Compressive strength is continuous
2. Independent variables should be categorical with two or more categories: Age group is categorical with 3 ordered categories
3. Observations should be independent: Yes they are independent.

4. There should be no significant outliers
```{r}
# Create a box plot to check for outliers
df %>% 
  ggplot(aes(x=age_group, y=strength, fill=age_group))+
  geom_boxplot()+
  labs(title = "Compressive Strength by Age group", y = "Compressive Strength (MPa)") +
  theme_bw()
```
There are some outliers present and they seem significant and not due to errors.


5. Dependent variable should be approximately normally distributed for each category of the independent variable
```{r}
df %>% 
  ggplot(aes(x=strength))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, alpha = 0.5, color='black', fill='steelblue')+
  geom_density(color='darkred')+
  labs(title = "Distribution of Concrete Compressive Strength across Age groups")+
  facet_grid(~age_group)+
  theme_bw()
```
The distribution of early age and middle age seem right skewed, mature age seems normal, but we can confirm with shapiro test
```{r}
byf.shapiro(strength ~ age_group, data=df)
```
All the p-values are less than 0.05, so none of them are normally distributed.

Let us check the normality of the transformed variable

```{r}
# test for normality on sqrt strength
byf.shapiro(sqrt_strength ~ age_group, data=df)
```
Two of the p-values are still less than 0.05 so they are not normal.

Since at least 2 of our assumptions were not met, we will use a non parametric test.


Assumptions of Kruskal Wallis test
1. Independence of Observations: Each observation is independent.
2. Ordinal or Continuous Outcome: Compressive strength is continuous.
3. Non-overlapping Groups: Each sample belongs to only one age group.
4. Similar Shape of Distributions: There appears to be some minor differences in their distributions.
Therefore we will also run the dunns test.


Kruskal-Wallis test
```{r}
# Run the Kruskal-Wallis test
kruskal.test(strength ~ age_group, data=df)
```
The p-value < 0.05, we reject the null hypothesis, indicating that compressive strength distributions differ significantly across the age groups.


3b. Post Hoc Test (Dunn Test) to see which specific age groups differ significantly from each other
```{r}
# run the Dunns test
dunnTest(df$strength, df$age_group, method='bonferroni')
```

The p-values are all less than 0.05, therefore, the Dunn’s test results with Bonferroni correction indicate significant differences in compressive strength across each of the three age groups.





4. Hypothesis Test 4: Difference in Compressive Strength Based on Water Content Levels

Null Hypothesis (H0): There is no difference in compressive strength among groups with different levels of water content.

Alternative Hypothesis (H1): There is a significant difference in compressive strength among groups with different levels of water content

```{r}
# Divide water into 3 groups based on quantiles
one_third <- quantile(df$water, 0.33)
two_third <- quantile(df$water, 0.66)
print(c(one_third, two_third))
```
so we divide the levels as:
water level < 172.34,  172.34 < water level < 192, and water level >= 192
```{r}
# create a new water_level column
df <- df %>% 
  mutate(water_level = case_when(
    water < one_third ~ "Low",
    water >= one_third & water < two_third ~ "Middle",
    water >= two_third ~ "High"
  ),
  water_level = factor(water_level, levels = c("Low", "Middle", "High"), ordered = TRUE))

head(df)
```

One way ANOVA ASSUMPTIONS
1. Dependent variable should be continuous: Compressive strength is continuous
2. Independent variables should be categorical with two or more categories: Water_level is categorical with 3 ordered categories
3. Observations should be independent: Yes they are independent.

4. There should be no significant outliers
```{r}
# Create a box plot to check for outliers
df %>% 
  ggplot(aes(x=water_level, y=strength, fill=water_level))+
  geom_boxplot()+
  labs(title = "Compressive Strength by Water Level", y = "Compressive Strength (MPa)") +
  theme_bw()
```
There are some outliers, but it doesn't seem to be influential so we can approve the 4th assumption.

5. Dependent variable should be approximately normally distributed for each category of the independent variable
```{r}
# check distributions for normality
df %>% 
  ggplot(aes(x=strength))+
  geom_histogram(aes(y = after_stat(density)), bins = 10, alpha = 0.5, color='black', fill='steelblue')+
  geom_density(color='darkred')+
  labs(title = "Distribution of Concrete Compressive Strength across Water Levels")+
  facet_grid(~water_level)+
  theme_bw()
```
```{r}
# plot qq plots to check for normality
df %>% 
  ggplot(aes(sample=strength))+
  stat_qq_point(size = 2,color = "steelblue")+
  stat_qq_line(color="darkred")+
  labs(title = "QQ Plot for Water Level", x = 'Theoretical', y = 'Sample')+
  facet_grid(~water_level)+
  theme_bw()
```
The QQ plots show that all three levels are approximately normally distributed.

6. Variances of the dependent variable within each category should be homogeneous
```{r}
# use bartlett.test
bartlett.test(strength ~ water_level, data=df)
```
the p-value is less than 0.05, indicating evidence that the variances are not equal.

We can run a WELCH ANOVA instead.

```{r}
# Run the WELCH ANOVA test
oneway.test(strength ~ water_level, data=df, var.equal = FALSE)
```
p-value < 0.05, We reject the null hypothesis, indicating a statistically significant difference in compressive strength across the different water content groups (Low, Medium, and High). We can conclude that at least one of the water levels is different from the others in terms of compressive strength.

In order to find out which levels are different, we run a post hoc test


4b. Post Hoc Test (Tukey HSD Test) to see which specific water levels differ significantly from each other
```{r}
# run anova using aov function
water_aov <- aov(strength ~ water_level, data=df)
# run Tukey HSD test
water_post_test <- TukeyHSD(water_aov) 
water_post_test
```
The TUKEY HSD test results show that all three adjusted p-values are less than 0.05, so we reject the null hypothesis, which means that all water levels are significantly different in terms of compressive strength.
```{r}
par(mar = c(5, 5, 5, 5))
plot(water_post_test)
```


```{r}
```


